Methods for Large Scale Learning to Rank

### 2014-11-20 report/Chapters/Conclusions.tex
kleine fixes thesis

To answer this research question, we took the approach of implementing the list of algorithms found in answering the first research question, starting with the Learning to Rank method with the highest certainty on ranking accuracy, ListNet.\\

>>>>

To answer this research question, we implemented the ListNet algorithm that was found to have the highest certainty in ranking accuracy.\\

### 2014-11-20 report/Chapters/Future.tex
kleine fixes thesis

\section{Distributed Computing Models}
We explored the possibilities of Hadoop MapReduce for distributed computation of Learning to Rank training algorithms. Hadoop, since the introduction of Hadoop YARN in Hadoop 2.0, offers integration with other distributed computing models, including Dryad \cite{Isard2007}, Spark \cite{Zaharia2010}, Storm \cite{Aniello2013} and \ac{MPI}. Of these distributed programming models, Spark is particularly promising, since Shukla et al. \cite{Shukla2012} already showed that it is able to speed-up the ListNet training procedure with much lower job scheduling overhead than that we found for MapReduce. For the newly supported programming models in Hadoop, it holds that they lack the critical mass as the distributed programming model of choice. This lack of a critical mass results in higher integration costs and less available support. The integration of new programming model into the Hadoop YARN framework is a step in the good direction to alleviate the integration cost of these programming models, but does not yet completely eliminate the problem. Cloud-based Hadoop services like Microsoft HDInsight and Amazon Elastic MapReduce.

\section{Learning to Rank Algorithms}
As an answer to the first research question of this thesis, we found five Learning to Rank methods for it holds that no other algorithm produced more accurate rankings with a higher degree of certainty of ranking accuracy. These algorithms were, from highest ranking accuracy / lowest certainty to lowest ranking accuracy / highest certainty: ListNet, SmoothRank, FenchelRank, FSMRank and LRUF. In this thesis we explored the speed-up characteristics of ListNet on Hadoop MapReduce. The speed-up characteristics of SmoothRank, FenchelRank, FSMRank and LRUF are using the MapReduce computing model are still to be explored.

@@

>>>>

\section{Learning to Rank Algorithms}
As an answer to the first research question of this thesis, we found five Learning to Rank methods for it holds that no other algorithm produced more accurate rankings with a higher degree of certainty of ranking accuracy. These algorithms were, from highest ranking accuracy / lowest certainty to lowest ranking accuracy / highest certainty: ListNet, SmoothRank, FenchelRank, FSMRank and LRUF. In this thesis we explored the speed-up characteristics of ListNet on Hadoop MapReduce. The speed-up characteristics of SmoothRank, FenchelRank, FSMRank and LRUF when implemented using the MapReduce computing model are still to be explored.

@@

\section{Distributed Computing Models}
We explored the possibilities of Hadoop MapReduce for distributed computation of Learning to Rank training algorithms. Hadoop, since the introduction of Hadoop YARN in Hadoop 2.0, offers integration with other distributed computing models, including but not limited to Dryad \cite{Isard2007}, Spark \cite{Zaharia2010}, Storm \cite{Aniello2013} and \ac{MPI}. Of these distributed programming models, Spark is particularly promising, since Shukla et al. \cite{Shukla2012} already showed that it is able to speed-up the ListNet training procedure with much lower job scheduling overhead than that we found for MapReduce. For the newly supported programming models in Hadoop, it holds that they lack the critical mass as the distributed programming model of choice. This lack of a critical mass results in higher integration costs and less available support. The integration of new programming model into the Hadoop YARN framework is a step in the good direction to alleviate the integration cost of these programming models, but does not yet completely eliminate the problem. The rise of cloud-based Hadoop services like Microsoft HDInsight and Amazon Elastic MapReduce and their increasing support for programming models other than MapReduce further alleviates these integration cost, as it takes away the need for high expenses on hardware.

### 2014-11-20 report/Chapters/Results.tex
laatste punten Dolf :):):)
kleine fixes thesis

\subsection{Speedup}
\label{ssec:speedup}
Figure \ref{fig:listnet_train_time} (linear data size axis) and Figure \ref{fig:listnet_train_time_log} (logarithmic data size axis) show the processing times of a single iteration of the ListNet training algorithm, using the ListNet implementation included in the RankLib library as well as the cluster implementation described in chapter \ref{chap:implementation} with different numbers of data nodes. The horizontal positions of the measurements are identical between execution method, as they are originate from the data sizes of the data sets used. Measurements were performed on a set of data sets consisting of the LETOR 3.0 OHSUMED data set, LETOR 4.0 data sets and the MSLR-WEB10/30K data sets are used, supplemented with generated data sets that are multiplications of MSLR-WEB30K (as described in section \ref{ssec:rq2_methodology}). Table \ref{tbl:recap_datasets} describes the data sets used in the experiments and their single training fold data sizes.\\

@@

@@

Figure \ref{fig:listnet_processing_speed} shows the processing speed in bytes per second. Our observation that RankLib ListNet processes very slow for data sets that do not fit in physical memory and virtual memory is needed is very notable in this graph. Furthermore, this graph shows how both increase in number of data nodes in a cluster and increase in input data size result in an increase in processing speed. The sub-linear growth of processing speed as a factor of input data size can again be explained as a result of the non-parallelisable Hadoop job scheduling part of the operation.\\

A result of that the Hadoop job scheduling overhead does not grow when the input data grows, is that the speed-up is linear or potentially better than linear as a function of data size. This can be seen in Figure \ref{fig:listnet_train_time}. In this Figure the cluster run lines are linear or, in case of the 24 and 32 data nodes cluster lines, slightly better than linear. The lines of the clusters of 8 data nodes or smaller have high variance, this can be attributed to the fact that these clusters were mostly tested on very small datasets, leaving the measurement very dependent on the variance of the job scheduling overhead.

@@

\section{SmoothRank}

>>>>

\subsection{Speedup}
\label{ssec:speedup}
Figure \ref{fig:listnet_train_time} (linear data size axis) and Figure \ref{fig:listnet_train_time_log} (logarithmic data size axis) show the processing times of a single iteration of the ListNet training algorithm, using the ListNet implementation included in the RankLib library as well as the cluster implementation described in chapter \ref{chap:implementation} with different numbers of data nodes. The horizontal positions of the measurements are identical between execution method, as they are originate from the data sizes of the data sets used. Measurements were performed on a set of data sets consisting of the LETOR 3.0 OHSUMED data set, LETOR 4.0 data sets and the MSLR-WEB10/30K data sets are used, supplemented with generated data sets that are duplications of MSLR-WEB30K (as described in section \ref{ssec:rq2_methodology}). Table \ref{tbl:recap_datasets} describes the data sets used in the experiments and their single training fold data sizes.\\

@@

The measurements of the clusters of 8 data nodes or smaller in Figure \ref{fig:listnet_train_time} have high variance, this can be attributed to the fact that these clusters were mostly tested on very small datasets, leaving the measurement very dependent on the variance of the job scheduling overhead. When we look at the more stable measurements, the ones on the clusters of 24 data nodes or more, we see that the processing time of a ListNet training iteration increases either linearly of very slightly sub-linearly as a function of the data size.\\

@@

Figure \ref{fig:listnet_processing_speed} shows the processing speed in bytes per second. Our observation that RankLib ListNet processes very slow for data sets that do not fit in physical memory and virtual memory is needed is very notable in this graph. This graph shows how both an increase in number of data nodes in a cluster and an increase in input data size result in an increase in processing speed. The super-linear increase of processing speed as a function of data size that can be seen in the 2, 4 and 8 data-node-cluster-lines can be explained as a result of the non-parallelisable Hadoop job scheduling part of the operation. For the lines originating from the 24, 32 and 64 data node clusters, we see a slowing growth of the processing speed as a function of data size. Note that the processing speed of the clusters is bounded by their CPU processing speed and will therefore not increase indefinitely as a function of data size.\\

@@
